### 2019-06-11 doc/chapters/towards_verifying_addition.tex
typo
fixed title

After realizing, that it would consume to much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, the smallest unit in Bitcoin-S-Core that is worthwhile to verify was extracted.
This could be the addition of two CurrencyUnits.
To make it even easier, the addition of CurrencyUnits with zero.
CurrencyUnits is an abstract class in Bitcoin-S, representing currencies like Satoshis.

@@

The additions signature looks like this:
\begin{lstlisting}[language=scala]
  +(c: CurrencyUnit): CurrencyUnit
\end{lstlisting}

@@

Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and fixed it to Int64.
This limits the verification a bit, but as we only want to verify the addition in satoshis, thats OK.

@@

This is easy.
We just replace all occurence of A with BigInt, since A is not overrided in an implementing class.
This is not exact the same code, because an implementing class could not override A anymore but thats fine for our verification.

@@

Now we get som weird warnings about require.

\section{Removeing Second Parameter of require}

@@

Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
We can saftly remove it.

A new error apperas.

@@

This error is hard to understand but we can see that there is a missing Long to BigInt conversion.
So we search for all longs in the code.
\begin{lstlisting}[language=scala]
  private case class Int64Impl(underlying: BigInt) extends Int64 {
    require(underlying >= -9223372036854775808L)
    require(underlying <= 9223372036854775807L)
  }
\end{lstlisting}

@@

\section{Propagating require}

@@

Sure, when adding one to the maximum Int64 the require does not hold anymore.
Since we do only allow zero as a parameter the easiest way is to restrict it to zero here too.

>>>>

After realizing, that it would consume too much time to rewrite the Bitcoin-S code and even the extracted part with checkTransaction, the smallest unit in Bitcoin-S-Core that is worthwhile to verify was extracted.
This could be the addition of two CurrencyUnits.
To make it even easier, the addition of CurrencyUnits with zero.
CurrencyUnits is an abstract class in Bitcoin-S, representing currencies like Satoshis.

@@

The additions' signature looks like this:
\begin{lstlisting}[language=scala]
  +(c: CurrencyUnit): CurrencyUnit
\end{lstlisting}

@@

Since there is no easy way to fix it and the code should stay as much as possible the original, we just remove the abstract type and fixed it to Int64.
This limits the verification a bit, but as we only want to verify the addition in satoshis, that's OK.

@@

This is easy.
We just replace all occurrence of A with BigInt, since A is not overrode in an implementing class.
This is not exact the same code, because an implementing class could not override A anymore but that's fine for our verification.

@@

Now we get some weird warnings about require.

\section{Removing Second Parameter of Require}

@@

Seems like Stainless does not support the second string parameter of require or at least it throws a warning about it.
We can safely remove it.

A new error appears.

@@

This error is hard to understand, but we can see that there is a missing Long to BigInt conversion.
So we search for all longs in the code.
\begin{lstlisting}[language=scala]
  private case class Int64Impl(underlying: BigInt) extends Int64 {
    require(underlying >= -9223372036854775808L)
    require(underlying <= 9223372036854775807L)
  }
\end{lstlisting}

@@

\section{Propagating Require}

@@

Sure, when adding one to the maximum Int64 the require does not hold anymore.
Since we do only allow zero as a parameter, the easiest way is to restrict it to zero here too.

### 2019-06-11 doc/chapters/trying_verify_checktx.tex
typo

Some of the code in this section is copied or adapted from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
Bitcoin-S-Core has a bitcoin transaction builder class with the following constructor:
\begin{lstlisting}[language=scala]
  BitcoinTxBuilder(
    destinations: Seq[TransactionOutput], // where to send the money
    utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
    feeRate: FeeUnit,                     // fee rate per byte
    changeSPK: ScriptPubKey,              // where to send the change
    network: BitcoinNetwork               // bitcoin network information
  ): Future[BitcoinTxBuilder]
\end{lstlisting}

The return type Future does not make sense here, since the implementing constructor calls either Future.successfull or Future.fromTry which returns an already resolved Future.
This might be for future purposes.

@@

First, we need a some money.
Thus, we create a fake transaction with one single output.
This transaction can be parsed from the bitcoin network but we create one manually in order to see this process.
\begin{lstlisting}[language=scala]
  val privKey = ECPrivateKey.freshPrivateKey
  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)

@@

This utxs are the inputs of our transaction.

@@

Finally, we define the fee rate in satoshis per one byte transaction size as well as some bitcoin network parameters.
The bitcoin network parameters are not important so we use some static values for testing purpose.
\begin{lstlisting}[language=scala]
  val feeRate = SatoshisPerByte(Satoshis.one)

@@

What should not be possible is a transaction referencing the same output twice.
This bug occured in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018. \cite{cve201817144}

@@

This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019 through this work along a unit test to prevent this bug from appearing again in the future.

@@

After integrating Stainless in Bitcoin-S there are many errors.
It takes too much time to fix them all so it should be easier to extract the classes needed for checkTransaction.

The extracted code has more than 1500 lines.
After running Stainless on it, it still throws a huge bunch of errors about what Stainless can not reason about.
After fixing some of those errors there appear new ones.
So this would require to change nearly everything of the extracted code.

Lets adjust the property from the \nameref{property_1} to the \nameref{property_2}, because there is not enough time to fix all this errors and write the verification.
This is way less interesting to verify but gives still many insights in what needs to be done to verify some parts of Bitcoin-S and other source code.

So lets look at the \nameref{property_2}.

>>>>

Some code in this section is copied or adapted from Bitcoin-S-Core transaction builder example.\cite{BitcoinSCore:txbuilderexample}
Bitcoin-S-Core has a bitcoin transaction builder class with the following constructor:
\begin{lstlisting}[language=scala]
  BitcoinTxBuilder(
    destinations: Seq[TransactionOutput], // where to send the money
    utxos: BitcoinTxBuilder.UTXOMap,      // unspent transaction outputs
    feeRate: FeeUnit,                     // fee rate per byte
    changeSPK: ScriptPubKey,              // where to send the change
    network: BitcoinNetwork               // bitcoin network information
  ): Future[BitcoinTxBuilder]
\end{lstlisting}

The return type Future does not make sense here, since the implementing constructor calls either Future.successful or Future.fromTry which returns an already resolved Future.
This might be for future purposes.

@@

First, we need a some money.
Thus, we create a fake transaction with one single output.
This transaction can be parsed from the bitcoin network, but we create one manually in order to see this process.
\begin{lstlisting}[language=scala]
  val privKey = ECPrivateKey.freshPrivateKey
  val creditingSPK = P2PKHScriptPubKey(pubKey = privKey.publicKey)

@@

This utxos are the inputs of our transaction.

@@

Finally, we define the fee rate in satoshis per one byte transaction size as well as some bitcoin network parameters.
The bitcoin network parameters are not important, so we use some static values for testing purpose.
\begin{lstlisting}[language=scala]
  val feeRate = SatoshisPerByte(Satoshis.one)

@@

What should not be possible is a transaction referencing the same output twice.
This bug occurred in Bitcoin Core known as CVE-2018–17144 which was patched on September 18, 2018. \cite{cve201817144}

@@

This was fixed in \href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request \mypound435} on GitHub at April 23, 2019, through this work along a unit test to prevent this bug from appearing again in the future.

@@

After integrating Stainless in Bitcoin-S, there were many errors.
It takes too much time to fix them all so it should be easier to extract the classes needed for checkTransaction.

The extracted code has more than 1500 lines.
After running Stainless on it, it still throws a huge bunch of errors about what Stainless can not reason about.
After fixing some of those errors, there appear new ones.
So this would require changing nearly everything of the extracted code.

Let's adjust the property from the \nameref{property_1} to the \nameref{property_2}, because there is not enough time to fix all these errors and write the verification.
This is way less interesting to verify but gives still many insights in what needs to be done to verify some parts of Bitcoin-S and other source code.

So let's look at the \nameref{property_2}.
